---
title: "PPT Images"
author: "Lok Yiu Xuan [23171563]"
date: "`r format(Sys.Date(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(splines)
library(rms)
library(lme4)
library(ggplot2)
library(patchwork)
library(dplyr)
library(tidyverse)
library(gganimate)
source("../scripts/helpers.R")
theme_set(my_theme)
```

```{r}
set.seed(2026)

n_studies <- 4
participants <- c(10, 15, 12, 18)

traditional <- tibble(
  Study = paste0("Study ", 1:n_studies),
  Effect = c(0.5, 0.7, 0.6, 0.8),
  SE = c(0.1, 0.12, 0.09, 0.11)
)

ipd <- bind_rows(lapply(1:n_studies, function(i) {
  tibble(Study = paste0("Study ", i),
         Effect = rnorm(participants[i], mean = traditional$Effect[i], sd = traditional$SE[i]))
}))

pooled_trad <- sum(traditional$Effect / traditional$SE^2) / sum(1 / traditional$SE^2)
pooled_se_trad <- 1 / sqrt(sum(1 / traditional$SE^2))

pooled_ipd <- mean(ipd$Effect)
pooled_se_ipd <- sd(ipd$Effect) / sqrt(nrow(ipd))

y_limits <- round(c(min(traditional$Effect), max(traditional$Effect)), 1) + c(-0.1,0.1)

p_trad <-  ggplot(traditional, aes(Study, Effect)) +
  geom_point(aes(color = "Study mean"), size = 4) +
  geom_errorbar(aes(ymin = Effect - SE, ymax = Effect + SE, color = "Study mean"), width = 0.2, size = 1) +
  geom_hline(aes(yintercept = pooled_trad, color = "Pooled effect"),
             linetype = "dashed", size = 1.5) +
  geom_rect(xmin = 0.5, xmax = n_studies + 0.5,
            ymin = pooled_trad - 1.96*pooled_se_trad,
            ymax = pooled_trad + 1.96*pooled_se_trad,
            fill = "red", alpha = 0.05) +
  scale_color_manual(values = c("Study mean" = "darkblue",
                                "Pooled effect" = "red")) +
  labs(title = "Traditional Meta-Analysis",
       x = "Study", y = "Effect Size",
       color = "") +
  coord_cartesian(ylim = y_limits) +
  my_theme

p_ipd <- ggplot(ipd, aes(Study, Effect)) +
  geom_jitter(aes(color = "Participant"), width = 0.2, size = 2, alpha = 0.6) +
  geom_point(data = traditional, aes(Study, Effect, color = "Study mean"), size = 4) +
  geom_errorbar(data = traditional, aes(Study, ymin = Effect - SE, ymax = Effect + SE, color = "Study mean"), width = 0.2, size = 1) +
  geom_jitter(aes(Study, Effect, color = "Participant"),
              width = 0.2, size = 2, alpha = 0.6) +
  geom_hline(aes(yintercept = pooled_ipd, color = "Pooled effect"),
             linetype = "dashed", size = 1.5) +
  geom_rect(xmin = 0.5, xmax = n_studies + 0.5,
            ymin = pooled_ipd - 1.96*pooled_se_ipd,
            ymax = pooled_ipd + 1.96*pooled_se_ipd,
            fill = "red", alpha = 0.005) +
  scale_color_manual(values = c("Participant" = "forestgreen",
                                "Study mean" = "darkblue",
                                "Pooled effect" = "red")) +
  labs(title = "IPD Meta-Analysis",
       x = "Study", y = "Effect Size",
       color = "") +
  coord_cartesian(ylim = y_limits) +
  my_theme


p_trad
p_ipd
```


```{r}
library(ggplot2)
library(dplyr)
set.seed(2026)

# ----------------------------
# Simulate participant-level data
# ----------------------------
n_studies <- 4
participants_per_study <- c(10, 15, 12, 18)

ipd <- bind_rows(
  lapply(1:n_studies, function(i){
    tibble(
      Study = paste0("Study ", i),
      Participant = 1:participants_per_study[i],
      Effect = rnorm(participants_per_study[i], mean = 0.5 + 0.1*i, sd = 0.1)
    )
  })
)

# ----------------------------
# Study-level summary (mean ± SE)
# ----------------------------
study_summary <- ipd %>%
  group_by(Study) %>%
  dplyr::summarize(
    MeanEffect = mean(Effect),
    SE = sd(Effect)/sqrt(n()),
    .groups = "drop"
  )

# ----------------------------
# Pooled effect across studies
# ----------------------------
pooled_effect <- sum(study_summary$MeanEffect / study_summary$SE^2) /
                 sum(1 / study_summary$SE^2)
pooled_se <- 1 / sqrt(sum(1 / study_summary$SE^2))
# ----------------------------
# Final IPD figure
# ----------------------------
ggplot() +
  # Participant-level points
  geom_jitter(data = ipd, aes(x = Study, y = Effect),
              width = 0.2, height = 0, size = 2, color = "forestgreen", alpha = 0.6) +
  # Study-level summary (mean ± SE)
  geom_point(data = study_summary, aes(x = Study, y = MeanEffect),
             color = "darkblue", size = 4) +
  geom_errorbar(data = study_summary, aes(x = Study, ymin = MeanEffect - SE, ymax = MeanEffect + SE),
                width = 0.2, color = "darkblue", size = 1) +
  # Pooled estimate line
  geom_hline(yintercept = pooled_effect, color = "red", linetype = "dashed", size = 1.5) +
  geom_rect(aes(xmin = 0.5, xmax = n_studies + 0.5, ymin = pooled_effect - 1.96*pooled_se, ymax = pooled_effect + 1.96*pooled_se),
          fill = "red", alpha = 0.1)+
  labs(
    title = "IPD Meta-Analysis: Participant-Level and Study-Level Effects",
    subtitle = "Green dots = participants, Blue dot + error = study summary, Red dashed = pooled effect",
    x = "Study",
    y = "Effect Size"
  ) + my_theme
```


# Helper Functions

## Logistic Curve Generator
```{r}
generate_curve <- function(age, midpoint, slope, increasing = TRUE,
                           noise_sd = 0, min_value = 0.01, max_value = 0.99) {

  sign <- ifelse(increasing, -1, 1)

  y <- min_value + (max_value - min_value) /
    (1 + exp(sign * slope * (age - midpoint)))

  if (noise_sd > 0) y + rnorm(length(age), 0, noise_sd) else y
}
```

## Logistic Model Used in Simulation

```{r}
logistic <- function(age,
                     baseline = 10,
                     amplitude = 5,
                     slope = 0.15,
                     midpoint = 60) {

  baseline + amplitude / (1 + exp(slope * (age - midpoint)))
}
```

# Example S Shaped Dose Response Patterns

## Simulation Settings

```{r}
age_seq <- seq(20, 80, length.out = 200)
n_points <- 40

curve_types <- list(
  "Increasing (e.g., mortality risk)" =
    list(slope = 0.09, midpoint = 55, noise_sd = 0.05, increasing = TRUE),

  "Decreasing (e.g., physical function)" =
    list(slope = 0.15, midpoint = 40, noise_sd = 0.08, increasing = FALSE)
)
```

## Generate Curves

```{r}
df_curves <- do.call(rbind, lapply(names(curve_types), function(type_name) {

  p <- curve_types[[type_name]]

  data.frame(
    Age = age_seq,
    Response = generate_curve(age_seq, p$midpoint, p$slope, p$increasing),
    Type = type_name
  )
}))
```

## Generate Noisy Points

```{r}
df_points <- do.call(rbind, lapply(names(curve_types), function(type_name) {

  p <- curve_types[[type_name]]
  age_values <- runif(n_points, 25, 75)
  data.frame(
    Age = age_values,
    Response = generate_curve(
      age_values,
      p$midpoint,
      p$slope,
      p$increasing,
      p$noise_sd
    ),
    Type = type_name
  )
}))
```

## Plot

```{r}
ggplot() +
  geom_point(data=df_points,
             aes(Age, Response, color = Type),
             alpha = 0.4,
             size = 1.5,
             position = position_jitter(width = 0.5)) +

  geom_line(data=df_curves,
            aes(Age, Response, color = Type),
            size = 1.2) +

  labs(
    title = "S Shaped Dose Response Patterns",
    x = "Age",
    y = "Response"
  ) +

  my_theme +
  scale_color_manual(values = c("darkred", "navy")) +
  theme(legend.position = "bottom", legend.title = element_blank())
```

# Multi Study Simulation Patterns

## Age Generation with Coverage Gaps

```{r}
generate_age_points <- function(n, min_age, max_age, gaps = NULL) {

  if (is.null(gaps)) {
    return(sort(runif(n, min_age, max_age)))
  }

  segments <- cbind(
    c(min_age, gaps + 5),
    c(gaps, max_age)
  )

  lens <- segments[,2] - segments[,1]

  n_seg <- round(n * lens / sum(lens))

  ages <- unlist(
    mapply(runif, n_seg, segments[,1], segments[,2], SIMPLIFY = FALSE)
  )

  sort(ages)[1:n]
}
```

## Study Specifications

```{r}
n_points <- 30

curve_types <- list(

  "Study 1" = list(
    slope = 0.25,
    midpoint = 42,
    noise_sd = 0.06,
    increasing = FALSE,
    min_age = 25,
    max_age = 75,
    gaps = NULL,
    color = "darkred",
    line_type = "solid"
  ),

  "Study 2" = list(
    slope = 0.12,
    midpoint = 55,
    noise_sd = 0.04,
    increasing = FALSE,
    min_age = 45,
    max_age = 70,
    gaps = NULL,
    color = "#ff7f0e",
    line_type = "dashed"
  ),

  "Study 3" = list(
    slope = 0.18,
    midpoint = 50,
    noise_sd = 0.07,
    increasing = FALSE,
    min_age = 30,
    max_age = 70,
    gaps = c(45,55),
    color = "#2ca02c",
    line_type = "dotted"
  )
)
```

## Generate Study Curves

```{r}
df_curves <- do.call(rbind, lapply(names(curve_types), function(type_name) {

  p <- curve_types[[type_name]]

  data.frame(
    Age = age_seq,
    Response = generate_curve(age_seq, p$midpoint, p$slope, p$increasing),
    Type = type_name,
    LineType = p$line_type
  )
}))
```

## Generate Study Points

```{r}
df_points <- do.call(rbind, lapply(names(curve_types), function(type_name) {

  p <- curve_types[[type_name]]

  age_points <- generate_age_points(
    n_points,
    p$min_age,
    p$max_age,
    p$gaps
  )

  data.frame(
    Age = age_points,
    Response = generate_curve(
      age_points,
      p$midpoint,
      p$slope,
      p$increasing,
      p$noise_sd
    ),
    Type = type_name
  )
}))
```

## Plot Multi Study Curves

```{r}
colors <- sapply(curve_types, `[[`, "color")
linetypes <- sapply(curve_types, `[[`, "line_type")

ggplot() +

  geom_point(data=df_points,
             aes(Age, Response, color = Type),
             alpha = 0.5,
             size = 1.8,
             position = position_jitter(width = 0.8)) +

  geom_line(data=df_curves,
            aes(Age, Response, color = Type, linetype = Type),
            size = 1) +

  labs(
    title = "S Shaped Patterns Across Different Studies",
    x = "Age",
    y = "Response"
  ) +

  my_theme +
  scale_color_manual(values = colors) +
  scale_linetype_manual(values = linetypes) +
  theme(legend.position = "bottom", legend.title = element_blank())
```

# Logistic Parameter Sensitivity Animations

These animations illustrate how individual parameters affect the logistic curve.

## Age Grid

```{r}
age <- seq(20, 80, length.out = 200)
```

## Slope Animation

```{r}
df_slope <- expand_grid(
  Age = age,
  slope = seq(0.05, 0.30, length.out = 40)
) %>%
  mutate(Y = logistic(Age, slope = slope))

p_slope <- ggplot(df_slope, aes(Age, Y)) +
  geom_line(size = 1.2, color = "navy") +
  labs(
    title = "Slope: Y = 10 + 5 / (1 + exp({round(frame_time,2)}(Age - 60)))",
    x = "Age",
    y = "Response"
  ) +
  my_theme

animate(
  p_slope + transition_time(slope),
  nframes = 40,
  fps = 20,
  width = 450,
  height = 300,
  renderer = file_renderer("logit_slope/", prefix = "", overwrite = TRUE)
)

# animate(p_slope + transition_time(slope))
# anim_save("logit_slope.gif  ")
```

## Midpoint Animation

```{r}
df_mid <- expand_grid(
  Age = age,
  midpoint = seq(40, 70, length.out = 40)
) %>%
  mutate(Y = logistic(Age, midpoint = midpoint))

p_mid <- ggplot(df_mid, aes(Age, Y)) +
  geom_vline(aes(xintercept = midpoint), 
             linetype = "dashed", 
             color = "red", 
             alpha = 0.7) +
  geom_line(size = 1.2, color = "navy") +
  labs(
    title = "Mid Point: Y = 10 + 5 / (1 + exp(0.1({round(frame_time,2)} - 60)))",
    x = "Age",
    y = "Response"
  ) +
  my_theme

animate(
  p_mid + transition_time(midpoint),
  nframes = 40,
  fps = 20,
  width = 450,
  height = 300,
  renderer = file_renderer("logit_midpoint/", prefix = "", overwrite = TRUE)
)

# animate(p_mid + transition_time(midpoint))
# anim_save("logit_mid.gif")
```

## Amplitude Animation

```{r}
df_amp <- expand_grid(
  Age = age,
  amplitude = seq(2, 8, length.out = 40)
) %>%
  mutate(Y = logistic(Age, amplitude = amplitude))

p_amp <- ggplot(df_amp, aes(Age, Y)) +
  geom_ribbon(aes(ymin = 10, ymax = Y), alpha = 0.1, fill = "red") +
  geom_line(size = 1.2, color = "navy") +
  labs(
    title = "Amplitude: Y = 10 + {round(frame_time,1)} / (1 + exp(0.1(Age - 50)))",
    x = "Age",
    y = "Response"
  ) +
  my_theme

animate(
  p_amp + transition_time(amplitude),
  nframes = 40,
  fps = 20,
  width = 450,
  height = 320,
  renderer = file_renderer("logit_amplitude/", prefix = "", overwrite = TRUE)
)

# animate(p_amp + transition_time(amplitude))
# anim_save("logit_amplitude.gif")
```

## Baseline Animation

```{r}
df_base <- expand_grid(
  Age = age,
  baseline = seq(5, 15, length.out = 40)
) %>%
  mutate(Y = logistic(Age, baseline = baseline))

p_base <- ggplot(df_base, aes(Age, Y)) +
  geom_hline(aes(yintercept = baseline), linetype = "dashed", alpha = 0.3, col="red") +
  geom_line(size = 1.2, color = "navy") +
  labs(
    title = "Baseline: Y = {round(frame_time,1)} + 5 / (1 + exp(0.1(Age - 50)))",
    x = "Age",
    y = "Response"
  ) +
  my_theme


animate(
  p_base + transition_time(baseline),
  nframes = 40,
  fps = 20,
  width = 450,
  height = 300,
  renderer = file_renderer("logit_baseline/", prefix = "", overwrite = TRUE)
)

# animate(p_base + transition_time(baseline))
# anim_save("logit_baseline.gif")
```
